### 2020-04-20 final_report/data.tex
Complete draft of dimensionality reduction: pca and variograms

@@

@@

@@

@@

>>>>

\newpage

@@

Model training over high-dimensional parameter spaces may be facilitated by carefully reducing the number of variables used to describe the space. For many applications, feature selection strategies succeed in identifying a sufficiently representative subset of the original input variables; however, all given variables were assumed to be physically relevant to the MC TBR model. Feature extraction methods, on the other hand, aim to identify a transformation from the parameter space which decreases dimensionality; even if no individual parameter is separable from the space, some linear combinations of parameters or nonlinear functions of parameters may be.

\begin{figure}[h]
  \centering
    \includegraphics[width=0.6\linewidth]{fig2_pca.jpg}
    \caption{Cumulative variance for optimal features identified by PCA}
  \label{fig:pca}
\end{figure}

\begin{wrapfigure}{r}{0.5\textwidth}
  \vspace{-55pt}
  \begin{center}
    \hspace*{-.3\columnsep}\includegraphics[width=0.58\textwidth]{fig3_allvar.jpg}
    \caption{Semivariograms for MC TBR data with coolant materials: (a) $He$, (b) $H_2O$, (c) $D_2O$}
    \label{fig:var}
  \end{center}
  \vspace{-110pt}
\end{wrapfigure}

@@

To pursue linear feature extraction, principal component analysis (PCA) was performed via scikit-learn on a set of 300,000 uniform samples of the MC TBR model.

Figure \ref{fig:pca} shows the resultant cumulative variance of the feature vectors identified by PCA. The similar share of variance among all eleven features demonstrated irreducibility of the TBR model by linear methods.

@@

Kriging, or Gaussian process regression, is a geostatistical surrogate modelling technique which relies on correlation functions over distance (lag) in the paramater space [2]. Although kriging performed poorly for our use case due to high dimensionality, these correlation measures gave insight into similarities between discrete-parameter slices of the data.

Figure \ref{fig:var} shows the Matheron semivariance [3] for three discrete slices with coolant material varied, but all other discrete parameters fixed. Fits [4] to the Matérn covariance model confirmed numerically that the coolant material is the discrete parameter with the greatest distinguishability in the MC TBR model.

** Make sure discrete slices are mentioned before this **

@@

Question: is it accurate to say this represents nonlinear feature extraction? re my intro paragraph

\newpage

### 2020-04-20 final_report/introduction.tex
Complete draft of dimensionality reduction: pca and variograms

Instead, modern D-T reactors rely on tritium breeding blankets, specialised layers of material which partially line the reactor and produce tritium upon neutron bombardment, e.g. by

\begin{equation}
    ^1_0n + \hspace{3pt} ^6_3Li \rightarrow \hspace{3pt} ^3_1T + \hspace{3pt} ^4_2He
\end{equation}
\begin{equation}
    ^1_0n + \hspace{3pt} ^7_3Li  \rightarrow \hspace{3pt} ^3_1T + \hspace{3pt} ^4_2He + \hspace{3pt} ^1_0n
\end{equation}
\begin{wrapfigure}{r}{0.4\textwidth}
  \begin{center}
    \includegraphics[width=0.48\textwidth]{fig1_tokamakdiagram.png}
  \end{center}
  \caption{Typical single-null reactor configuration as specified by BLUEPRINT [6]: 1 — plasma,
2 — breeding blankets }
\end{wrapfigure}

where T represents tritium and $^7Li$, $^6Li$ are the more and less frequently occurring isotopes of lithium, respectively. $^6Li$ has the greatest tritium breeding cross-section of all tested isotopes [4], but due to magnetohydrodynamic instability of liquid lithium in the reactor environment, a variety of solid lithium compounds are preferred.

The TBR is defined as the ratio between tritium generation in the breeding blanket per unit time and tritium fuel consumption in the reactor. The MC neutronics simulations previously mentioned therefore must account for both the internal plasma dynamics of the fusion reactor and the resultant interactions of neutrons with breeding blanket materials. Neutron paths are traced through a CAD model (e.g. Figure 1) of a reactor with modifiable geometry.

The input parameters of the computationally-expensive TBR model therefore fall into two classes. Continuous parameters, including material thicknesses and packing ratios, describe the geometry of a given reactor configuration. Discrete categorical parameters further specify all relevant material sections, including coolants, armours, and neutron multipliers. One notable exception is the enrichment ratio denoting the presence of $^6Li$, a continuous parameter. Our challenge, put simply, was to produce a TBR function which takes these same input parameters and approximates the MC TBR model with the greatest achievable accuracy.

>>>>

Instead, modern D-T reactors rely on tritium breeding blankets, specialised layers of material which partially line the reactor and produce tritium upon neutron bombardment, e.g. by 
\begin{align}
    ^1_0n + \hspace{3pt} ^6_3Li &\rightarrow \hspace{3pt} ^3_1T + \hspace{3pt} ^4_2He \\
    ^1_0n + \hspace{3pt} ^7_3Li &\rightarrow \hspace{3pt} ^3_1T + \hspace{3pt} ^4_2He + \hspace{3pt} ^1_0n
\end{align}
\begin{wrapfigure}{r}{0.45\textwidth}
  \vspace{-20pt}
  \begin{center}
    \includegraphics[width=0.48\textwidth]{fig1_tokamakdiagram.png}
  \end{center}
  \caption{Typical single-null reactor configuration as specified by BLUEPRINT [6]: 1 — plasma,
2 — breeding blankets }
\label{fig:tokamak}
\end{wrapfigure}
where T represents tritium and $^7Li$, $^6Li$ are the more and less frequently occurring isotopes of lithium, respectively. $^6Li$ has the greatest tritium breeding cross-section of all tested isotopes [4], but due to magnetohydrodynamic instability of liquid lithium in the reactor environment, a variety of solid lithium compounds are preferred.

The TBR is defined as the ratio between tritium generation in the breeding blanket per unit time and tritium fuel consumption in the reactor. The MC neutronics simulations previously mentioned therefore must account for both the internal plasma dynamics of the fusion reactor and the resultant interactions of neutrons with breeding blanket materials. Neutron paths are traced through a CAD model (e.g. Figure \ref{fig:tokamak}) of a reactor with modifiable geometry.

The input parameters of the computationally-expensive TBR model therefore fall into two classes. Continuous parameters, including material thicknesses and packing ratios, describe the geometry of a given reactor configuration. Discrete categorical parameters further specify all relevant material sections, including coolants, armours, and neutron multipliers. One notable exception is the enrichment ratio, a continuous parameter denoting the presence of $^6Li$. Our challenge, put simply, was to produce a TBR function which takes these same input parameters and approximates the MC TBR model with the greatest achievable accuracy.
